---
title: Network analyses
output: html_document
---

### 1. Install required libraries

```{r}
# install.packages("remotes")
# remotes::install_github("zdk123/SpiecEasi")
# if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
# install.packages("GGally")
```

### 2. Load required libraries

```{r}
library(phyloseq)
library(SpiecEasi)
library(igraph)
library(intergraph)
library(network)
library(GGally)
```

### 3. Load R objects from diversity analyses

```{r}
load("../02-diversity_analyses/.RData")
# OR load data from current folder
# load(".RData")
```

### 4. ASV interaction network

#### 4a. First subset phyloseq object into different tooth type / HIV status groups

```{r}
ps.dat.noUS <- subset_samples(ps.dat, study_group != "US")
# subset phyloseq object into different groups
## HI CF-PF
ps.dat.HI.CFPF <- subset_samples(ps.dat.noUS, study_group == "HI" & aliquot_type == "CF-PF")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HI.CFPF <- filter_taxa(ps.dat.HI.CFPF, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                              
## HI CA-PD                              
ps.dat.HI.CAPD <- subset_samples(ps.dat.noUS, study_group == "HI" & aliquot_type == "CA-PD")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HI.CAPD <- filter_taxa(ps.dat.HI.CAPD , function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                              
## HUU CF-PF                              
ps.dat.HUU.CFPF <- subset_samples(ps.dat.noUS, study_group == "HUU" & aliquot_type == "CF-PF")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HUU.CFPF <- filter_taxa(ps.dat.HUU.CFPF, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                               
## HUU CA-PD                               
ps.dat.HUU.CAPD <- subset_samples(ps.dat.noUS, study_group == "HUU" & aliquot_type == "CA-PD")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HUU.CAPD <- filter_taxa(ps.dat.HUU.CAPD, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                               
## HEU CF-PF                               
ps.dat.HEU.CFPF <- subset_samples(ps.dat.noUS, study_group == "HEU" & aliquot_type == "CF-PF")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HEU.CFPF <- filter_taxa(ps.dat.HEU.CFPF, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                               
## HEU CA-PD                               
ps.dat.HEU.CAPD <- subset_samples(ps.dat.noUS, study_group == "HEU" & aliquot_type == "CA-PD")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HEU.CAPD <- filter_taxa(ps.dat.HEU.CAPD, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                               
## HI CA-PE                               
ps.dat.HI.CAPE <- subset_samples(ps.dat.noUS, study_group == "HI" & aliquot_type == "CA-PE")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HI.CAPE <- filter_taxa(ps.dat.HI.CAPE, function(x) sum(x > 3) > (0.10*length(x)), TRUE)

## HUU CA-PE                              
ps.dat.HUU.CAPE <- subset_samples(ps.dat.noUS, study_group == "HUU" & aliquot_type == "CA-PE")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HUU.CAPE <- filter_taxa(ps.dat.HUU.CAPE, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
                               
## HEU CA-PE                               
ps.dat.HEU.CAPE <- subset_samples(ps.dat.noUS, study_group == "HEU" & aliquot_type == "CA-PE")
# filter low abundance taxa to simplify network building step (seen at least 3 times in 10% of samples)
ps.dat.HEU.CAPE <- filter_taxa(ps.dat.HEU.CAPE, function(x) sum(x > 3) > (0.10*length(x)), TRUE)
```

#### 4b. HI CF-PF network

```{r}
# make image folder
system("mkdir img")
spiec.out <- spiec.easi(ps.dat.HI.CFPF, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HI.CFPF)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HI.CFPF)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HI.CFPF, type='taxa', color="V5")
grph
pdf("img/network.HI.CFPF.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
```

Visualize network with positive and negative edges

```{r}
getrank <- "V5"
# Col vector up to 74 color samples
col_vector74 = c("#7FC97F","#BEAED4","#FDC086","#FFFF99","#386CB0","#F0027F","#BF5B17","#666666","#1B9E77","#D95F02","#7570B3","#E7298A","#66A61E","#E6AB02","#A6761D","#666666","#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928","#FBB4AE","#B3CDE3","#CCEBC5","#DECBE4","#FED9A6","#FFFFCC","#E5D8BD","#FDDAEC","#F2F2F2","#B3E2CD","#FDCDAC","#CBD5E8","#F4CAE4","#E6F5C9","#FFF2AE","#F1E2CC","#CCCCCC","#E41A1C","#377EB8","#4DAF4A","#984EA3","#FF7F00","#FFFF33","#A65628","#F781BF","#999999","#66C2A5","#FC8D62","#8DA0CB","#E78AC3","#A6D854","#FFD92F","#E5C494","#B3B3B3","#8DD3C7","#FFFFB3","#BEBADA","#FB8072","#80B1D3","#FDB462","#B3DE69","#FCCDE5","#D9D9D9","#BC80BD","#CCEBC5","#FFED6F")
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HI.CFPF_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HI.CFPF.txt", format="ncol")
```

K-core decomposition

```{r}
# which ASVs are the core of the network 
which(coreness(spiec.graph)>=2)
```

```{r}
# periphery of the network
which(coreness(spiec.graph)==0)
```

Eigenvector centrality

```{r}
# measure of being well connected to the well connected -- influence of a node in a network - high eigenvector score means that a node is connected to many nodes who themselves have high scores
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HICFPF <- spiec.graph
```

#### 4c. HUU CF-PF network

```{r}
spiec.out <- spiec.easi(ps.dat.HUU.CFPF, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HUU.CFPF)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HUU.CFPF)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HUU.CFPF, type='taxa', color="V5")
grph
pdf("img/network.HUU.CFPF.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
```

Visualize network with positive and negative edges

```{r}
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HUU.CFPF_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HUU.CFPF.txt", format="ncol")
```

K-core decomposition

```{r}
# which ASVs are the core of the network 
which(coreness(spiec.graph)>=2)
```

```{r}
# periphery of the network
which(coreness(spiec.graph)==0)
```

Eigenvector centrality

```{r}
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HUUCFPF <- spiec.graph
```

#### 4d. HEU CF-PF network

```{r}
spiec.out <- spiec.easi(ps.dat.HEU.CFPF, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HEU.CFPF)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HEU.CFPF)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HEU.CFPF, type='taxa', color="V5")
grph
pdf("img/network.HEU.CFPF.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
```

Visualize network with positive and negative edges

```{r}
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HEU.CFPF_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HEU.CFPF.txt", format="ncol")
```

K-core decomposition

```{r}
# which ASVs are the core of the network 
which(coreness(spiec.graph)>=2)
```

```{r}
# periphery of the network
which(coreness(spiec.graph)==0)
```

Eigenvector centrality

```{r}
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HEUCFPF <- spiec.graph
```

#### 4e. HI CA-PD network

```{r}
spiec.out <- spiec.easi(ps.dat.HI.CAPD, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HI.CAPD)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HI.CAPD)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HI.CAPD, type='taxa', color="V5")
grph
pdf("img/network.HI.CAPD.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HI.CAPD_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HI.CAPD.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HICAPD <- spiec.graph
```

#### 4f. HUU CA-PD network

```{r}
spiec.out <- spiec.easi(ps.dat.HUU.CAPD, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HUU.CAPD)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HUU.CAPD)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HUU.CAPD, type='taxa', color="V5")
grph
pdf("img/network.HUU.CAPD.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HUU.CAPD_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HUU.CAPD.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HUUCAPD <- spiec.graph
```

#### 4g. HEU CA-PD network

```{r}
spiec.out <- spiec.easi(ps.dat.HEU.CAPD, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HEU.CAPD)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HEU.CAPD)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HEU.CAPD, type='taxa', color="V5")
grph
pdf("img/network.HEU.CAPD.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HEU.CAPD_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HEU.CAPD.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HEUCAPD <- spiec.graph
```

#### 4g. HI CA-PE

```{r}
spiec.out <- spiec.easi(ps.dat.HI.CAPE, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HI.CAPE)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HI.CAPE)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HI.CAPE, type='taxa', color="V5")
grph
pdf("img/network.HI.CAPE.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HI.CAPE_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HI.CAPE.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HICAPE <- spiec.graph
```

#### 4h. HUU CA-PE

```{r}
spiec.out <- spiec.easi(ps.dat.HUU.CAPE, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HUU.CAPE)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HUU.CAPD)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HUU.CAPE, type='taxa', color="V5")
grph
pdf("img/network.HUU.CAPE.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HUU.CAPE_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HUU.CAPE.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HUUCAPE <- spiec.graph
```

#### 4i. HEU CA-PE network

```{r}
spiec.out <- spiec.easi(ps.dat.HEU.CAPE, method="mb", lambda.min.ratio=1e-2, nlambda=20, pulsar.params=list(rep.num=100))
spiec.graph <- adj2igraph(getRefit(spiec.out), vertex.attr=list(name=taxa_names(ps.dat.HEU.CAPE)))
paste("Total number of ASVs in graph: ", length(taxa_names(ps.dat.HEU.CAPD)))
# save network
grph <- plot_network(spiec.graph, ps.dat.HEU.CAPE, type='taxa', color="V5")
grph
pdf("img/network.HEU.CAPE.pdf")
grph
dev.off()
#how many positive and negative edges inferred from network?
betaMat <- as.matrix(symBeta(getOptBeta(spiec.out)))
positive <- length(betaMat[betaMat>0])/2
paste("number of positive connections: ", positive)
negative <- length(betaMat[betaMat<0])/2
paste("number of negative connections: ", negative)
total <- length(betaMat[betaMat!=0])/2
paste("total number of connections: ", total)
# visualize network with pos and neg connections
asv.ids <- colnames(spiec.out[[1]]$data)
edges <- E(spiec.graph)
edge_cols <- ifelse(betaMat>0, 'black', 'red')[upper.tri(betaMat) & betaMat!=0]
E(spiec.graph)$color <- edge_cols

#How many nodes connected at specific rank
nb_nodes <- vcount(spiec.graph)
tax_table(ps.dat) <- tax_table(ps.dat)[,getrank]
asv_ids <- V(spiec.graph)$name
idx <- which(row.names(tax_table(ps.dat)) %in% asv_ids)
taxa <- as.character(tax_table(ps.dat)[,getrank])[idx]
ig2 <- asNetwork(spiec.graph)
network.vertex.names(ig2) <- taxa
net <- ig2
net %v% getrank <- as.character(taxa)
y <- col_vector74[1:nb_nodes]
names(y) <- levels(as.factor(taxa))

#Plot the network
grph <- ggnet2(net, color = getrank, alpha=0.75, palette = y, size = 6, edge.size=1, edge.color="color", edge.alpha = 0.5, label = asv_ids, label.size = 2) +theme(legend.position = "none")
grph
pdf("img/network.HEU.CAPE_posneg.pdf")
grph
dev.off()

#write spiec-easi graph to file 
write.graph(spiec.graph, file="spieceasi.ncol.HEU.CAPE.txt", format="ncol")

# kcore decomp
paste("core network ASVs: ", which(coreness(spiec.graph)>=2))
paste("peripheral network ASVs: ", which(coreness(spiec.graph)==0))

# eigenvalue centrality
tail(sort(igraph::eigen_centrality(spiec.graph)$vector), 10)
# save graph
spiec.graph.HEUCAPE <- spiec.graph
```

Save workspace

```{r}
save.image()
```

